---
title: "Msnbase ggplot2"
author: "Jan Stanstrup"
date: "31 jul 2017"
output: html_document
---



# Get some test data
## Libraries
```{r, message=FALSE, warning=FALSE, echo = -1}
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
library(MSnbase)
```

## Get Chromatograms object using the MSnbase package
Load 6 of the CDF files from the faahKO
```{r}
cdf_files <- dir(system.file("cdf", package = "faahKO"), recursive = TRUE,
		 full.names = TRUE)[c(1:3, 7:9)]
```

<br>
Define the sample grouping.
```{r}
s_groups <- rep("KO", length(cdf_files))
s_groups[grep(cdf_files, pattern = "WT")] <- "WT"

s_groups
```

<br>
Define a data.frame that will be used as phenodata
```{r}
pheno <- data.frame(sample_name = sub(basename(cdf_files), pattern = ".CDF",
				      replacement = "", fixed = TRUE),
		    sample_group = s_groups, stringsAsFactors = FALSE)

pheno
```

<br>
Read the data and extract chromatograms
```{r}

raw_data <- readMSData2(cdf_files, pdata = new("NAnnotatedDataFrame", pheno)) 

bpis <- chromatogram(raw_data, aggregationFun = "max")

bpis
```


<br><br>

# Current baseplot

# some different ways to define the colors we will be using
```{r}
library(RColorBrewer)
sample_colors <- brewer.pal(3, "Set1")[1:2]
names(sample_colors) <- c("KO", "WT")

sample_colors
```


```{r}
col <- paste0(sample_colors[raw_data$sample_group], "80")
col
```

The "base" plot using the plotting feature from MSnbase.

```{r}
plot(bpis, col = col)
```






# Functions to do the plotting with ggplot2
## Functions to get chromatograms in long format for ggplot2
```{r}
.ChromatogramLongFormat <- function(x){
        data.frame(rtime = x@rtime, intensity = x@intensity, fromFile = x@fromFile, row.names = NULL, stringsAsFactors = FALSE)
}
```
        
```{r}
.ChromatogramsLongFormat <- function(x, pdata = NULL){
        out <- lapply(x@.Data, .ChromatogramLongFormat)
        out <- do.call(rbind.data.frame,out)
        
        # if we supplied pdata add the corresponding annotation to each row
        if(!is.null(pdata)){
         pdata <- cbind.data.frame(fromFile =  1:nrow(pdata), pdata@data, stringsAsFactors = FALSE)
         out <- merge(out, pdata, all.y = TRUE,  by = "fromFile")
        }
        
        return(out)
}
```

## function that make the basic ggplot
```{r}
ggplotChromatograms <- function(bpis, rt_unit = "sec", pdata = NULL) {
        require(ggplot2)
        bpis_long <- .ChromatogramsLongFormat(bpis, pdata = pdata)
        
        if(rt_unit=="min"){
                bpis_long$rtime <- bpis_long$rtime/60
                x_lab <- "Retention time (min)"
        }
        
        if(rt_unit=="sec"){
                x_lab <- "Retention time (sec)"
        }
        
        ggplot(data = bpis_long, aes(x = rtime, y = intensity, group = fromFile)) + 
                geom_line() +
                theme_classic() +
                labs(x = x_lab, y = "Intensity")
}
```



# User function in action
## Basic plot
```{r}
ggplotChromatograms(bpis)
```

## Use minutes instead
```{r}
p <- ggplotChromatograms(bpis, rt_unit = "min")
p
```



## color by sample...
... assigning a new color to each sample using default ggplot2 color
```{r}
p + aes(color=as.factor(fromFile))
```

## define color for each file manually
This does the same as the original base plot

```{r}
names(col) <- 1:length(col)

p + aes(color=as.factor(fromFile)) +
    scale_colour_manual(name = "File", values = col)
```

## Use injected sample info
We can instead add some info about each sample
We can then use the columns in this table to modify the plot
```{r}


p <- ggplotChromatograms(bpis, rt_unit = "min", pdata = new("NAnnotatedDataFrame", pheno)) +
aes(color=sample_group)

p
```

## Manually define colors
```{r}
p + scale_colour_manual(name = "Group", values = sample_colors)
```




# TODO
* handle several mz-ranges

